Quantifying Stability and Change in Personal Culture Using Panel Data

### 2023-09-20 manuscript.Rmd
results and discussion

A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and nondeclarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].

@@

The pattern of results help clarify and, hopefully, transcend previous debates. We observe similar patterns across the data sets that have produced conflicting interpretations. In all datasets we see similar rates of individual-level change over time across a broad range of measures of personal culture. But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines. Whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time. Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults.

@@

We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016]. The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel. The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences.

@@

Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel. All panels show a similar range of the proportion of systematic variance attributable to intrapersonal change, with all panels including questions that show essentially no systematic variance attributable to intrapersonal change and all panels including a few questions where more than 20 percent of systematic variance is attributable to intrapersonal change. Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses. Again, this is to be expected. Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture. Despite the differences in interpretation present in previous works, results from the GSS are highly consistent with results from the other panels.

PARAGRAPH HERE ABOUT RESULTS: MOST/LEAST SYSTEMATIC VARIANCE. DISTRIBUTION. NEED THAT INFO.

Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed. In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 50 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 0 and 15 percent of total variance. Residual variance tends to account for between 20 to 40 percent of variance (though on several questions residual variance is greater than 50 percent).

The substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question. However, we observe a mean of about 9.7 percent of explained variance being attributed to intrapersonal change across all the questions. If we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences.

@@

Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features. We find that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change. We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change. The earlier a question was asked, even net of how long it has been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.

The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question. Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change. The fact that we find a negative coefficient attached to this variable –-the longer we observe people, the less variance is explained by intrapersonal change-– is a challenge to that assumption.

As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question. If the coefficient reflects a true negative "causal" effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions. Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model. The longer we observe a question, the higher proportion of systematic variance is attributable to intrapersonal change. We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.

@@

While these broad results help clarify previous debates, they mostly reframe previous empirical results. To better illustrate the value of our approach, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.

@@

This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, and preferences that manifest at the individual level. Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate. We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people. On some questions (such as factors that affect one’s life satisfaction or attitudes about government benefits), it is highly plausible that changes in personal culture as a result of peoples’ experiences as adults are the primary explanation for differences in personal culture.

However, the amount of intrapersonal change is often small relative to interpersonal differences (approximately 10 percent across all surveys). On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero. These include questions about civil liberties, abortion, interindividual (kk note: generalized?) trust, or sense of civic duty. On these questions, assuming the window of observation is not uniquely stable, the amount of variance attributable to intrapersonal change does not seem sufficient to explain how and why adults end up with different personal cultures. For these items, the explanation for personal cultural differences seems to reside principally in early-life experiences.

It is difficult to draw a single conclusion about the relative importance of intrapersonal change in explaining cultural differences across a wide range of questions measured on different scales and over different durations of time. A key assertion of this work is that blanket statements of “change” versus “stability” as the single pattern among adults is likely to be insufficient. Change and stability are always present among adults, and no viable theoretical model expects just one process to explain all difference. But the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys. There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.

To be clear, this does not mean that adults do not change or are incapable of changing. On basically all questions, some members of the sample change their personal culture. And our results should not be taken to imply that intrapersonal change is not important or consequential. Small numbers of changers can have important consequences in the aggregate. A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences. Our results could also be interpreted as suggesting that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon. Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.

@@

While we find that people continually reshape their personal culture over time, our results continue to direct attention to the “past conditions of production” that shape peoples’ personal culture early in life. To the extent that we wish to understand what makes adults different, focusing attention on peoples’ transition between social roles, selection into and out of organizational environments, and restructuring of social networks, as well as changes in the cultural resources people have at their disposal, seem to explain a much smaller proportion of their differences than the formative experiences that happen early in life.

The findings we present here are consonant with a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018]. While some people clearly change over time as they move into new roles, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles. Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.

This stability in adulthood could be driven by a range of theoretical processes. It could be the case that adults are exposed to new information at comparable rates as young people but are resistant to changing their personal culture. It could also mean that they are just given fewer opportunities to change their personal culture. For example, adult social environments like workplaces and voluntary associations might expose people to less new information than adolescent social environments such as schools and colleges.

@@

Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven by cohort replacement –- one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population [@vaisey2016]. This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.

@@

Second, our results continue to highlight the value of panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone. The process of how individual people store and report personal culture is key to understand the aggregate processes of social change.

Additionally, we echo previous calls for studies –-including panel studies-– of the socialization processes of young people. Our results suggest that much of the work of social differentiation takes place before people turn 18. In order to understand why adults are different, we need to understand them before they are adults.

@@

The approach outlined here generates estimates of how much variance in a population is due to each of the two mechanisms outlined here. It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes. Adjudicating this question will require different models that distinguish different kinds of change.

Similarly, the models presented here, even as they allow different slopes for different respondents, still assume that change is a linear function of time. This assumption likely does not hold true in practice. Change over time likely takes non-linear forms. People might jump from one “stable” disposition to another, for example. This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases.

Similarly, our approach, in assuming a linear trajectory for change, assumes that durable change is unidirectional. To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition, this is modeled as residual within-person variance, rather than one of our two theoretical processes.

## Implications for Future Research

We believe our results push past the “needless dichotomy” of asking whether people change their personal culture established by previous work. Our central hope is that, in quantifying the level of intrapersonal change compared to the amount of interpersonal difference at baseline, researchers can start focusing on the individual and social conditions that affect the amount of change we observe.

For example, our results do not speak to the specific mechanisms that limit change in personal culture during adulthood. It is highly plausible that we observe limited amounts of intrapersonal change because the structuring of adulthood rarely provides the social circumstances necessary for intrapersonal change to happen, and that a different social structuring of adulthood would produce higher rates of intrapersonal change. Rather than adults being resistant to change, it could be the case that they are shielded from opportunities for change.

Researchers with strong prior expectations about the social factors that matter for the relative importance or intrapersonal change can partition samples and test hypotheses about the relative importance of intrapersonal change across these partitions.

# Conclusion

In this paper we sought to move past the “needless dichotomy” implicit in the question of whether people change or not. Accepting that some amount of change in personal culture will appear when explored over a long enough time scale, at a high enough degree of resolution, or across a large number of questions, we sought to quantify the relative contribution of intrapersonal change and interpersonal difference across a large range of personal culture, to make broader claims about the importance of change to cultural differences.

Our findings presented here should not be construed as suggesting that social experiences during adulthood are irrelevant for personal culture or can never have meaningful effects. Clearly identifying causal effects of experiences, social transitions, or interventions on personal culture is an important undertaking, even if those treatments do not explain a large variance in personal culture across people. It is only through such studies that researchers will be able to better clarify the mechanisms that produce personal cultural change.

Our hope is that the findings presented here clarify what are actually highly compatible findings from recent debates. People do change. Just not very much.

>>>>

A recurring focus in the sociology of culture is the question of whether people, as they move through their lives and transition between social roles, change their personal culture –- the declarative and non-declarative attitudes, beliefs, values, and dispositions that manifest at the individual level [@lizardo2017; @kiley2020; @lersch2023].

@@

The pattern of results help clarify and, hopefully, transcend previous debates. We observe similar patterns across the data sets that have produced conflicting interpretations. In all data sets we see similar rates of individual-level change over time across a broad range of measures of personal culture. But quantifying these processes also reveals that intrapersonal change plays a substantially smaller role in explaining the variance in peoples’ personal culture than between-person differences in their baselines. Whether this amount of intrapersonal change is “meaningful” ultimately remains a theoretical question reflected in prior expectations concerning the question, population, and time. Nevertheless, the results suggest that the intrapersonal change observed (in some cases over more than 10 years) does not seem to play a substantial role in explaining personal cultural differences between adults.

@@

We can think of this number as akin to an Intraclass Correlation Coefficient focused only on the systematic components of the model while ignoring residual variance that might be attributable to measurement error or non-persisting change [see also the variance decomposition of period and cohort by @vaisey2016]. The higher the number is, the more variance in responses is attributable to intrapersonal change during the panel. The lower the number, the less intrapersonal change seems to play a large role in explaining cultural differences during the panel.

@@

Figure 1 plots the proportion of systematic variance attributable to interpersonal differences (“between”) and intrapersonal change (“within”) for each of the 609 questions analyzed, plotted separately by panel. To broadly summarize the results, all panels show a similar range of the proportion of systematic variance attributable to intrapersonal change. Across panels, the median proportion of systematic variance attributable to intrapersonal change is 0.094 (mean of 0.098), with an interquartile range of 0.053 to 0.139. All panels include questions that show essentially no systematic variance attributable to intrapersonal change and all panels include questions where more than 20 percent of systematic variance is attributable to intrapersonal change.

Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change. Again, this is to be expected. Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.

To the extent that there are differences across the panels, the PSID has the highest mean (.120) and median (.135) proportion of systematic variance attributable to intrapersonal change. While we cannot disentangle features of the sample from features of the questions asked to each sample, the specific samples for many PSID questions have lower average ages than those from other panels. To the extent that younger respondents might be more likely to make durable changes of opinion, this higher explanatory power of intrapersonal change might reflect the distinct age profile of respondents in this sample. At the other end, the GSS has the lowest mean (.073) and median (.069) proportion of variance attributable to intrapersonal change. This potentially reflects the fact that the GSS observes people for a shorter duration, on average, than the other panels. If, consistent with life course adaption theories, people are more likely to make significant cultural changes the longer we observe them, then duration likely affects the amount of variance attributable to intrapersonal change (a point we explore more below). However, the GSS results are still highly consistent with results from the other panels.

While there are some differences between panels, these differences are small compared to the differences within panels. For about 6 percent of items explored here, more than 20 percent of systematic variance is attributable to intrapersonal change. These questions tend to ask about external referents (e.g., confidence in specific government leaders or political parties), life satisfaction, or current financial position. At the other end, questions about religious identification, views on gender roles, and support for civil liberties tend to have very low proportion of variance attributable to intrapersonal change.

In contrast to the tournament of models approach, quantifying change this way allows us to explore variation in the relative importance of intrapersonal change across questions that all both show evidence of change. For example, Kiley and Vaisey [-@kileyvaisey2020] found that confidence in the press and confidence in religious leaders were both characterized by active updating, our results show that intrapersonal change is much more important for explaining variance in confidence in the press (0.164) than confidence in religion (0.049).

Appendix A presents the estimated proportion of variance attributable to interpersonal differences and intrapersonal change, as well as the proportion of residual variance, for each question analyzed. In general, interpersonal differences are almost always the largest component of the total variance and tend to explain between 55 and 70 percent of total variance, while intrapersonal change is always the smallest, typically explaining between 3 and 8 percent of total variance. Residual variance tends to account for between 22 and 37 percent of variance (though on several questions residual variance is greater than 50 percent).

As noted above, the substantive importance of intrapersonal change for understanding cultural differences depends on a range of factors, including how long the panel runs, whether assumptions about linear change hold, and whether the panel is capturing a distinctly turbulent period or a distinctly stable period for the item under question. However, if we assume that the period under observation is “typical” for a question –-not a time of extremely heightened (or lowered) sensitivity or change-– then it does not seem realistic for this process of intrapersonal change to explain a large share of interpersonal differences for many of the questions we observe.

@@

Figure 2 plots the results from an ordinary least squares regression of proportion of systematic variance attributable to intrapersonal change as a function of question, panel, and sample features. These models also include fixed effects for panels (not shown), so coefficients reflect the association between a particular sample or question characteristic within a panel.[^paneleffects]

[^paneleffects]: KK NOTE: Can we summarize these panel fixed effects at all?

Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the proportion of systematic variance is attributable to intrapersonal change. We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change. The earlier a question was asked, even net of how long it was been asked, the lower the proportion of systematic variance is attributable to intrapersonal change.

The most notable result in the meta analysis is the negative coefficient attached to the span of years covered by the question. The longer a question is observed, the less systematic variance is attributable to intrapersonal change. Theories of personal cultural change that link changes in personal culture to social experiences, including the LCAM, suggest that the longer we observe respondents, the more likely people are to undergo potentially transformative experiences and therefore the more variance would be explained by intrapersonal change. The fact that we find a negative coefficient attached to this variable is a challenge to that assumption.

As a further test of this coefficient, we compared the proportion of systematic variance attributable to intrapersonal change when using the full duration of a panel compared to when when we dropped the final wave and therefore reduced the total duration of observation for the question. If the coefficient reflects a true negative effect of duration on systematic variance attributable to intrapersonal change, we should see that same effect within questions. Results from this analysis are presented in Appendix B and contradict the coefficient from the regression model. The longer we observe the same question, the higher proportion of systematic variance is attributable to intrapersonal change. We interpret this combination of findings as suggesting that the kinds of questions asked for a longer time period tend to demonstrate less intrapersonal change than questions asked over shorter periods, rather than a true function of time.

This reinforces the explanation for why the GSS displays the lowest average proportion of systematic variance attributable to intrapersonal change: its shorter duration. [KK NOTE: Does this inform any other panel differences? Does the PSID, BHPS or UKHLS have longer durations on average?]

@@

While these broad results help speak to previous debates, they mostly re-frame previous findings without extending the theoretical debate. To better illustrate the value of our approach for addressing different theoretical questions, we turn to a specific empirical example: the relative importance of intrapersonal change and interpersonal difference for explaining cultural differences by education levels.

@@

This paper set out to clarify previous findings about the prevalence of durable change in personal culture -- the attitudes, beliefs, values, behaviors, and preferences that manifest at the individual level. Relying on an approach that quantified the amount of systematic variance attributable to either intrapersonal change or interpersonal differences at baseline, we found a consistent pattern across data sets previously used in this debate. We see evidence of intrapersonal change over time on almost all questions, with some measures of personal culture displaying quite high levels of intrapersonal change relative to the amount of variance between people. On some questions, such as the factors that affect one’s life satisfaction or attitudes about government benefits, it is highly plausible that differences in peoples’ experiences as adults are the primary explanation for differences in personal culture.

However, the amount of intrapersonal change is often small relative to interpersonal differences (less than 10 percent on average across all surveys). On many questions, the amount of systematic variance attributable to intrapersonal change is essentially zero. These include questions about civil liberties, abortion, inter-individual (kk note: generalized?) trust, or sense of civic duty. On these questions and the large number of others with a low proportion of systematic variance attributable to intrapersonal change, there does not appear to be sufficient cultural change during adulthood for experiences and transitions between social locations to be the primary explanation for why adults end up with different personal cultures. For these items, the explanation for personal cultural differences seems to reside principally in early-life experiences.

More important than adjudicating past debates, however, we show that the approach developed in this paper can adjudicate novel theoretical questions about the conditions in which change or stability is more prevalent and more important for explaining cultural differences within a population. As an example, we showed that the role of intrapersonal change in explaining cultural differences varies significantly by question and by education. Intrapersonal change plays almost no role in explaining general sentiments about the government's role in the economy, while it plays a much larger role in explaining ideological and partisan identification. At the same time, across issues, intrapersonal change is much more important for explaining cultural variation among people without a college degree than for people with a college degree. While this pattern might be confounded by age and other factors that affect college completion, it potentially suggests that college-going locks in personal culture in a way that other experiences do not, making adult experiences less influential for political dispositions for college graduates compared to the rest of the population.

To this point, the theoretical debate in cultural sociology has focused on the existence of personal cultural change among adults and its relative importance. We believe the "existence" debate is theoretically misguided. Personal cultural change will always exist to some degree in an adult population, and no viable theoretical model expects just one process --stability or change-- to characterize a population. Our results reinforce that assertion, finding a mixture of interpersonal differences and intrapersonal change for all items. Even when we observe very low levels of systematic variance attributable to intrapersonal change for particular questions, this is potentially attributable to insufficient data, insufficient resolution on the underlying belief, or insufficient duration of observation.

A more reasonable theoretical debate pushed by previous work is the relative importance of these two processes for explaining cultural variation. While change might always be prevalent, its impact on social differences might be small or large, and our approach has been designed to clarify the contribution of intrapersonal change to such differences. It is difficult to draw a single conclusion on this question across a wide range of questions measured on different scales and over different durations of time, but the overall picture, with important caveats, suggests that intrapersonal change during adulthood is not the dominant explanation for cultural differences among adults for most items we measure in surveys. There simply is not enough evidence of intrapersonal change on these items relative to the amount of difference we observe across people.

To be clear, none of our results should be interpreted as suggesting that adults do not change or are incapable of changing. On all questions, some members of the sample change their personal culture at least a little bit. Our results also should not be taken to imply that intrapersonal change is not important or consequential. Small numbers of changers can have important consequences in the aggregate. A shift from 49 percent supporting gay marriage to 51 percent supporting it can have significant political ramifications, even as most respondents remain fixed in their preferences. And our results should not detract from work trying to estimate the mechanisms through which personal cultural differences --however small-- emerge among adults. Just because change is small does not mean the mechanisms that produce is are not important to understand.

Our results should be read principally as suggesting that if researchers want to explain why people differ in their personal culture, it is imperative to look at the conditions in which they were raised and the experiences that happened before they reached adulthood. To the extent that we want to explain what makes adults different, focusing attention on peoples’ transition between social roles, their selection into and out of organizational environments, and changes in their social networks, as much sociological work does, appears to explain a much smaller proportion of their differences than the formative experiences that happen early in life.

In this way, our findings echo a range of recent causal inference work that suggests selection effects, rather than treatment effects, seem to explain the majority of differences in personal culture between people in different social roles or positions [@campbell2016; @wodtke2018]. While some people clearly change over time as they move into new social roles or into new social environments, this amount of change does not appear to be sufficient to explain what are often substantial differences between people in different roles. Our results suggest that, to the extent we observe differences in personal culture between people in different social roles, such as parenthood, having children, education, or working in positions of authority [@longest2013], or between people in different occupations [@weeden2005], our prior expectation should be that this difference is selection, rather than treatment.

Our results do not speak to exactly why intrapersonal change lacks significant explanatory power in understanding culture. First, it could be the case that adults are exposed to opportunities for change at the same rate, but become less willing to change. Or it could be that opportunities for durable change or the social conditions that facilitate durable change are just extremely uncommon. Changes in social structures, the life course, or cultural expectations could make intrapersonal change a much more consequential explanation for cultural differences.

@@

@@

Because intrapersonal change in personal culture does not appear to be a driving force in cultural differences among adults, our results suggest that processes of social and cultural change at the aggregate level are likely driven more by cohort replacement, one group of people with certain cultural dispositions exiting the population and a different set of people with a potentially different set of cultural dispositions entering the population, than by contemporaneous social conditions [@vaisey2016; @underwood2022]. This is likely true for both large-scale change in societies over time as well as changes within organizations, political parties, occupations, and other social aggregates.

However, because our results suggest that the role of formative experiences compared to contemporaneous environments differs by a person's level of education, changes in educational attainment over time, as well as changes in the social structuring of the life course might be shifting the relative importance of these two factors over time. If we observe further differences across socio-demographic groups like education, social class, and race/ethnicity in the relative importance of these processes, this would argue for a greater incorporation the interaction of demographic processes with processes of individual cultural change into our explanations of aggregate cultural change.

@@

Our results continue to highlight the value of long-running panel surveys for adjudicating different theories of culture that cannot be evaluated with cross-sectional data alone. Across our results, residual fluctuations in responses account for more than four times as much variance, on average, as linear persistent change, meaning that most change in responses is not systematic. Under these circumstances, extending panels beyond two waves provides significantly more leverage to understand how and why personal culture changes.

Additionally, we echo previous calls for more studies –-including panel studies-– of the socialization processes of young people. Our results suggest that most of the variance between adults emerges before people turn 18 or relatively soon after. In limiting our empirical perspective to experiences that happen after that point, social scientists potentially focus on questions that, while important, lack the capacity to explain major cultural differences.

@@

The approach outlined here generates estimates of how much variance in a population is due to each of two broad processes: intrapersonal change and interpersonal differences at baseline. It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes. Adjudicating this question will require different models that distinguish different kinds of change.

Our results and interpretations also hinge on how we have defined change and the assumptions built into the Life Course Adaption Model. This model treats cultural trajectories as varying linear slopes for each respondent, assuming that change is a linear function of time. This assumption likely does not hold true in practice. Change over time likely takes non-linear and discontinuous forms. People might jump from one “stable” disposition to another, for example, or experience a discontinuous "turning point" that upsets a stable trajectory. This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases. Panels with more waves might allow for more flexible models of change that do not require this assumption of linearity.

Similarly, in assuming a linear trajectory for change, our approach assumes that durable change is unidirectional. To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes. This is a potentially reasonable assumption on short panels -- change that does not last two years is hard to classify as durable -- but it might not be true in longer panels. Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.

Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United States, and the United Kingdom). This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being. Our data does not cover some dimensions of culture that might be relevant for other aspects of sociological exploration (artistic tastes, leisure activities, and time use, for example).

# Implications for Future Research

We believe the approach outlined here can push past the “needless dichotomy” implicit in the question of whether people change or not and address more theoretically informed questions about the conditions under which the relative importance of intrapersonal change and interpersonal differences matter for explaining cultural differences. Characterizing questions as displaying change or not can only take researchers so far, but the question of why some questions demonstrate more change while others do not, or why some groups are characterized by more stability than others, has the potential to weigh in on a broader range of theoretical debates. We hope researchers move toward those questions.

While we focused on education as a determinant of the explanatory power of intrapersonal change, our approach can readily be extended to other socio-demographic categories that might affect intrapersonal change -- race/ethnicity, religiosity, social class, gender, and political affiliation, for example -- or their intersections. Researchers with different data might be able to evaluate the relative importance of intrapersonal change and interpersonal differences across contexts as well, either national contexts if the same questions are administered across countries or organizational contexts if data traces students within schools, workers within workplaces, or congregation members within congregations.

The wide variation across questions in the proportion of systematic variation attributable to intrapersonal change also calls for a more systematic exploration for why some forms of culture are characterized by stability while others are characterized by change, and what the consequences of these differences for the organization of beliefs in the public might be.
